Marlowe: implementing and analysing financial contracts on blockchain

### 2019-12-11 papers/marlowe2019/paper.tex
Shorten Section 4

In the next two sections we give an overview of some of the work we have carried out with the aim of helping users be more confident that the Marlowe contracts they write do what they expect them to do.
We look at two complementary approaches: static analysis of concrete contracts in this section,  and formal verification of general properties about Marlowe in the next.

For some properties, given a particular contract, a computer can decidably determine whether the contract satisfies the property and, in the cases when it does not, it can provide a counter-example.

For Turing-complete languages, this process is usually undecidable in the general case, and for Marlowe contracts it is often NP-complete in general. However, in practice, SMT (Satisfaction Modulo Theories) solvers are often able to solve moderately large instances of this problem reasonably efficiently.

@@

SBV provides a monad called \texttt{SBV} which encapsulates potentially unknown values to the program. A program that takes arguments wrapped with the \texttt{SBV} monad can be used with normal arguments and it can also be passed symbolic parameters when used as part of a property that is passed to the solver.
For example, the following function:

\begin{verbatim}
primeFactors :: SBV Integer -> SBV Integer -> SBV Bool
primeFactors x y = x .> 1 .&& y .> 1 .&& x * y .== 25723297
\end{verbatim}

\noindent
can be used both with normal integers and with symbolic parameters, that is, parameters whose concrete value is not specified and which the solver will try to replace with values that satisfy or falsify the property:

\begin{verbatim}
*Main> primeFactors 98 314
False
*Main> sat primeFactors
Satisfiable. Model:
    s0 = 6353 :: Integer
    s1 = 4049 :: Integer
*Main> primeFactors 6353 4049
True
\end{verbatim}

\subsubsection{Limitations}

At the time of writing, there are some types of functions and data-types that SBV cannot translate to \texttt{SMTLib}. In particular, there are two that affected the way in which we use SBV with Marlowe and that prevent us from using the same version of the semantics for symbolic execution and normal execution.

Firstly, calls to recursive functions that are unbounded cannot be translated. As part of the process of converting functions to \texttt{SMTLib} terms, SBV unfolds them into a finite data-structure. For this reason, if the function is recursive and is not bounded by a non-symbolic value (a value that is known at the time of translation), the translation process will run out of memory.

Secondly, to the best our knowledge, SBV does not currently directly provide support for representing custom data types symbolically, except for zero-ary datatypes, i.e: enumerations.

In the following section, we explain how we worked around these limitations and give an overview of how we use SBV to avoid runtime problems with Marlowe contracts.

@@

Marlowe semantics are designed to use types to prevent many non-sensical contracts from being written. But there are potential problems which are harder to detect until runtime, for example, whether there will be enough money to issue all the payments declared in the contract. At this point, it may already be too late to fix them, and particularly so in the case of blockchain. The Marlowe semantics represents these errors that can be found at runtime as \texttt{Warnings}.

@@

In order to achieve this, we implemented a symbolic version of the semantics that returns a list of the warnings produced by a trace (a list of transactions input to the contract):

@@

We can see that this version of the implementation takes two concrete parameters: \texttt{Bounds} (see \textit{Bounds for the state and the inputs} section below) and \texttt{Contract} (the contract we want to analyse); and two symbolic parameters: \texttt{SSlotNumber} (the initial slot number) and \texttt{SList NTransaction} (the input list of transactions); and it returns \texttt{SList NTransactionWarning} (a symbolic list of transaction warnings).

@@

As we mentioned before, SBV does not currently seem to support in general the use of custom datatypes. We do not need to represent the \texttt{Contract} type symbolically because we know its value at the time of analysis, but we still have a number of other types like \texttt{TransactionWarning} that are custom datatypes and we need to use.

Fortunately, SBV does support tuples and the \texttt{Either} type; combinations of those two give us essentially the same functionality as custom datatypes that are not recursive and have no type-parameters. We can represent all types that Marlowe requires as combinations of \texttt{Either} and tuples, with the exception of the \texttt{Contract} type, but we do not need a symbolic version of the \texttt{Contract} type.

@@

@@

In the non-symbolic semantics, each step typically takes a \texttt{Contract}, a \texttt{State}, and a set of \texttt{Input}s; and it returns an updated \texttt{State} and a remaining \texttt{Contract}.

This is a problem because values that rely on symbolic values have to be themselves symbolic. Moreover,  the resulting remaining \texttt{Contract} depends on the \texttt{Input}s and \texttt{State}, which are both symbolic.

@@

However, in both the input and the \texttt{State} record there are several lists (representing finite maps) that are not explicitly bounded in the implementation. Some parts of the semantics are bounded by the length of these lists (or maps), such as the implementation of \texttt{Close} that refunds any remaining money in the contract accounts.
In order for the symbolic implementation to be finite, we need to find a bound for the length of these lists and maps.

With the current semantics of Marlowe, all the lists/maps in \texttt{State} and input can be bounded when looking at a particular contract. In particular, we can find out quite easily, by looking at the contract, how many parties there are, how many choice identifiers are used, how many accounts are used, and how many let binding identifiers are used.

It is less obvious that we can also set a bound for the input \texttt{Transaction} list as the maximum number of nested \texttt{When}s in the contract (for all the execution paths) plus one. We also provide a proof in Isabelle that shows that any longer list of transactions will result in an error (we discuss this in Section~\ref{subsubsec:bound_max_transaction_number}).

>>>>

Marlowe semantics use types to prevent many non-sensical contracts from being written. But there are potential problems which are harder to detect until runtime, for example, whether there will be enough money to issue all the payments declared in the contract. At that point, it may already be too late to fix them, particularly in the case of blockchain.

Fortunately, in the case of Marlowe, a computer can decidably determine whether a particular contract satisfies certain property before executing it, and it can provide a counter-example when it does not.

@@

SBV provides a monad called \texttt{SBV}, a function can use parameters wrapped in this monad to represent symbolic values. Functions that take symbolic values can be used as properties and passed to the solver; the solver will then try to replace the symbolic values with concrete values that satisfy or falsify the property.

@@

Marlowe semantics represents errors that can be found at runtime as \texttt{Warnings}.

@@

This property is essentially a symbolic version of the semantics that returns a list of the warnings produced by a symbolic trace (a symbolic list of transactions input to the contract):

@@

@@

SBV does not currently seem to support in general the use of custom datatypes. Fortunately, SBV supports tuples and the \texttt{Either} type. We can represent all types that Marlowe requires as combinations of \texttt{Either} and tuples, with the exception of the \texttt{Contract} type, but we do not need a symbolic version of the \texttt{Contract} type because we know its value at the time of analysis.

@@

However, in both the input and the \texttt{State} record there are several lists (representing finite maps) that are not explicitly bounded in the implementation. Some parts of the semantics are bounded by the length of these lists (or maps), such as the implementation of \texttt{Close}.
In order for the symbolic implementation to be finite, we need to find a bound for the length of these lists or maps.

Fortunately, we can infer a bound for all this lists quite straightforwardly. The least obvious one is the length of the list of transactions; we discuss the proof for this bound in Section~\ref{subsubsec:bound_max_transaction_number}.

@@

Values that rely on symbolic values have to be themselves symbolic, and the continuation \texttt{Contract} after each step depends on the \texttt{Input}s and \texttt{State}, which are both symbolic. But having the contract as a symbolic parameter would be inconvenient since it is recursive, we know it in advance, and we use it to bound the execution of the symbolic semantics.

@@
